---
title: "DATA 607 Project 2"
author: "Sergio Ortega Cruz"
date: "October 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load the library
# Check if the package is installed. If not, install the package
if(!require('tidyr')) {
  install.packages('tidyr')
  library(tidyr)
}

# Check if the package is installed. If not, install the package
if(!require('dplyr')) {
  install.packages('dplyr')
  library(dplyr)
}

# Check if the package is installed. If not, install the package
if(!require('ggplot2')) {
  install.packages('ggplot2')
  library(ggplot2)
}
```

## Choose 3 Datasets and  execute the proposed transformation tasks described in the discussion forum


*1st Dataset Djocovich from Mikhail Groysman

All the source CSV files are hosted in Github

```{r dataset1}
url <- "https://raw.githubusercontent.com/sortega7878/DATA607Project2/master/Djocovich Mikhail Groysman.csv"
download_file <- "djocovich.csv" # name of the file on the local machine after download

# the file will be downloaded to the working directory
downloader::download(url, download_file)

# read the file into a vector
conn <- file(download_file, open="r")
dfdj <- readLines(conn, warn = FALSE)

close(conn)
djoco <- read.csv("djocovich.csv", header=T, stringsAsFactors = F)
djoco
```

Rename and clean column names and Fill empty Disciplines Rows
```{r empty}
library(plyr)
djoco <- rename(djoco, c("ï..Discipline"="Disciplines", "WR"="WinRatio"))
djoco
```

Fill the Empty rows where value Singles or Doubles are needed

```{r sindob}
djoco$Disciplines [2:7]<-paste("Singles")
djoco$Disciplines [8:14]<-paste("Doubles")
djoco
```
Remove totals


```{r totals}

djoco<-filter(djoco,(djoco$Type!="Total"))
djoco

```
Substitute "-" with 0's

```{r zeroing}

djoco %>% mutate_all(funs(ifelse(.=="-", "0", .)))
```
Change all dashes for zero's

The final Format is Tidy and ready for processing


*Second Dataset Poverty Rates by Chester Poon

```{r dataset2}
url <- "https://raw.githubusercontent.com/sortega7878/DATA607Project2/master/Poverty Chester Poon.csv"
download_file <- "poverty.csv" # name of the file on the local machine after download

# the file will be downloaded to the working directory
downloader::download(url, download_file)

# read the file into a vector
conn <- file(download_file, open="r")
dfdj <- readLines(conn, warn = FALSE)

close(conn)
poverty <- read.csv("poverty.csv", header=T, stringsAsFactors = F)
head(poverty)
```

Change "Dirty" Column names

```{r names4}
poverty <- rename(poverty, c("ï..Table.with.row.headings.in.column.A.and.column.headings.in.rows.6.to.9."="Heading"))
head(poverty)

```

Extract the exact subset of data we want to work with

```{r slice}
poverty<-slice(poverty,9:17)
poverty<-subset(poverty,select = c(Heading,X,X.1,X.5,X.6))
head(poverty)
```
Clean empty rows and fill the year where needed

```{r race}
#Filing Yeaqr Columns
library(stringr)
poverty$X<-2016
poverty$X.5<-2017
poverty3<- subset(poverty,X.1!="")
poverty3
```
In order to put consolidated columns will separate everything in two different datasets and bind them later

```{r names}
library (gtools)

df1<-data.frame("Race Sex"=poverty3$Heading, "Year"=poverty3$X, "Poor Number"=poverty3$X.1)
df2<-data.frame("Race Sex"=poverty3$Heading, "Year"=poverty3$X.5, "Poor Number"=poverty3$X.6)
df3<-smartbind(df1,df2)
df3


```
The task Was so separate the columns into Sex and Race however this seems to be a tidier format for analysis since alll the categorical varibales are in a single column and the transformation into different columns will aggregate just a lot of NA or zeros

```{r dataset3}
library(stringr)
url <- "https://raw.githubusercontent.com/sortega7878/DATA607Project2/master/messi stats Hans Angrantz.csv"
download_file <- "messi.csv" # name of the file on the local machine after download

# the file will be downloaded to the working directory
downloader::download(url, download_file)

# read the file into a vector
conn <- file(download_file, open="r")
dfdj <- readLines(conn, warn = FALSE)

close(conn)
messi <- read.csv("messi.csv", header=T, stringsAsFactors = F)
messi
```

1) Remove the header and the lines
2)Get Column Season, Champion League goal
3)Create a data frame
4)Process the summary of the data (mean, median, 1st quartile and 3rd quartile)
5) histogram and scatter plot

```{r head}

#The easiest way of tostart is slice all the data we don't need
#Removing unnecesary columns
#poverty<-slice(poverty,1:17)
#messi<-subset(messi,select = c(Season,X.3))
messi<- subset(messi,X.3!="")
messi<-slice(messi,2:16)

messi<-subset(messi,select = c(Season,X.3))
messi

```

Cleaning Season Data and change names of columns

```{r season}

messi$Season[grep("[:digits:]", messi$Season)]
messi$Season<-substring(messi$Season,1,4)
messi <- rename(messi, c("X.3"="Goals"))
messi

```

Changing format to Data Frame and generate statistical data as requested

```{r summaryplts}


messi2 = as.data.frame(sapply(messi, as.numeric))
summary(messi2)
```

Generation of plots as requested in the problem

```{r plot2}

ggplot(messi2, aes(x=Season, y=Goals)) +
    geom_bar(stat="identity")  


ggplot(messi2, aes(x=Season, y=Goals)) +
    geom_point(shape=4)  

```

